---
title: "An Analysis of Coronavirus"
author: "Jedidiah Harwood"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


# Abstract

# Introduction

  For the past two years, the corona virus pandemic has been a subject of vexation. A specific concern that has arisen throughout the pandemic, is whether or not certain geographical areas, are more prone to outbreaks than others, and whether government interventions to reduce corona virus infections were effective.  In this paper, we shall utilize techniques such as Analysis of Variance, Analysis of Covariance, and Bayesian Inference to determine whether or not certain geographical regions were more prone to corona virus outbreaks than others, and whether government efforts to reduce coronavirus infections were effective. 
  

# Exploratory Data Analysis

```{r Data Processing}
library(tidyverse)
library(gridExtra)
library(COVID19)
covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
covid1 <- covid %>% 
  filter(WHO_region != "Other") %>% mutate(WHO_region = fct_recode(WHO_region, "Eastern Mediterranean"="EMRO",
                                 "Europe" = "EURO",
                                 "Africa" = "AFRO",
                                 "Western Pacific" = "WPRO",
                                 "Americas"="AMRO",
                                 "South-East Asia" = "SEARO")) 
covid1 <- covid1[which(is.na(covid1) == F), ]

covid2 <- COVID19::covid19() 
covid <- left_join(covid1, covid2, by = c("Date_reported" = "date", "Country_code" = "iso_alpha_2")) %>% filter(!is.na(New_cases)) %>% group_by(WHO_region) 
```

```{r Table 1: Cases by Region}
knitr::kable(covid %>% summarise('New_Cases' = sum(New_cases)), caption = "Table 1: Number of New Cases by Region", col.names = c("Region", "New Cases"))
```



# Model

# Analysis

# Diagnostics/Sensitivity Analysis

# Conclusion


  